### 2025-02-27 paper/sections/Abstract.tex
solved second supervisor comments

\begin{abstract}
    This paper explores multilingual emotion classification across binary classification, intensity estimation, and cross-lingual detection tasks. To address linguistic variability and limited annotated data, we evaluate various deep learning approaches, including transformer-based embeddings and traditional classifiers. After extensive experimentation, language-specific embedding models were selected as the final approach, given their superior ability to capture linguistic and cultural nuances. Experiments on high- and low-resource languages demonstrate that this method significantly improves performance, achieving competitive macro-average F1 scores. Despite these advances, challenges remain due to limited annotated data in underrepresented languages and the complexity of nuanced emotional expressions. The study highlights the need for robust, language-aware emotion recognition systems and emphasizes future directions, including expanding multilingual datasets, refining models.
\end{abstract}

>>>>

\begin{abstract}
    This paper explores multilingual emotion classification across binary classification, intensity estimation, and cross-lingual detection tasks. To address linguistic variability and limited annotated data, we evaluate various deep learning approaches, including transformer-based embeddings and traditional classifiers. After extensive experimentation, language-specific embedding models were selected as the final approach, given their superior ability to capture linguistic and cultural nuances. Experiments on high- and low-resource languages demonstrate that this method significantly improves performance, achieving competitive macro-average F1 scores. Notably, in languages such as Tigrinya and Kinyarwanda for cross-lingual detection task, our approach achieved a second-place ranking, driven by the incorporation of advanced preprocessing techniques. Despite these advances, challenges remain due to limited annotated data in underrepresented languages and the complexity of nuanced emotional expressions. The study highlights the need for robust, language-aware emotion recognition systems and emphasizes future directions, including expanding multilingual datasets and refining models.
\end{abstract}

### 2025-02-27 paper/sections/Conclusion.tex
solved second supervisor comments

Nonetheless, certain limitations persist. The scarcity of annotated data in underrepresented languages and the inherent complexity of nuanced emotional expressions continue to impede model performance. Future research should focus on expanding multilingual datasets, refining preprocessing methods, and exploring novel architectures to enhance model generalization and adaptability. Additionally, fine-tuning multilingual models specifically for low-resource languages could significantly improve emotion detection accuracy in these contexts. Such efforts will be instrumental in advancing the field of multilingual emotion analysis and in developing more robust, language-aware emotion recognition systems.

>>>>

Nonetheless, current limitations in multilingual emotion analysis include the lack of annotated data for underrepresented languages and challenges in capturing nuanced emotional expressions, both of which hinder model performance. Future research should prioritize expanding multilingual datasets, improving preprocessing techniques, and developing new architectures to boost model generalization and adaptability. Fine-tuning models for low-resource languages could also enhance emotion detection accuracy, advancing the field and creating more effective, language-aware emotion recognition systems.

### 2025-02-27 paper/sections/ExperimentalSetup.tex
solved second supervisor comments

This section describes the experimental setup, providing details about data splits, preprocessing steps, hyperparameter tuning, computational resources, and the tools and libraries utilized. The goal is to ensure reproducibility, transparency, and alignment with established best practices in machine learning research.
\input{tables/MainResults}

@@

Hyperparameter tuning was conducted using \texttt{Optuna}, an advanced framework for automated hyperparameter optimization \citep{Akiba2019Optuna}. The hyperparameters tuned included the hyperparameters of SVM and XGBoost. A Bayesian optimization strategy was employed, balancing exploration of the hyperparameter space and exploitation of promising configurations. The performance of each hyperparameter configuration was evaluated on the validation set, and the best configuration was selected based on the performance metric of interest.

@@

The model fine-tuning with LoRA, and the training of the MLP and XGBoost models, utilized Binary Cross-Entropy (BCE) as the loss function for Tracks A and C, and Cross-Entropy for Track B, owing to its appropriateness for classification tasks. Meanwhile, the training of the SVM model employed hinge loss.Using LoRA, we fine-tuned the Q, K, and V matrices for feature extractor transformer models, as shown in Table \ref{tab:model_selection}. Given the unbalanced dataset, a weighted loss approach was employed to ensure that the model adequately learned from all classes. Optimization for fine-tuning deep learning models was performed using the \texttt{AdamW} optimizer, which improves upon the standard Adam optimizer by decoupling weight decay and learning rate updates \citep{Loshchilov2019AdamW}. To further enhance training stability and convergence, a cosine annealing learning rate scheduler with restarts\cite{loshchilov2017sgdrstochasticgradientdescent} was employed. This scheduling approach helped adaptively reduce the learning rate over time, facilitating better exploration of the loss landscape and improving generalization. The model was trained for a fixed number of epochs, and early stopping was used to terminate training if the validation performance plateaued, thus avoiding overfitting.

@@

The implementation of the experiments utilized several state-of-the-art tools and libraries. The deep learning models were implemented and trained using \texttt{PyTorch}, which provides a flexible and high-performance framework for machine learning research \citep{Paszke2019PyTorch}. For data manipulation and evaluation metrics, \texttt{Scikit-Learn} was employed \citep{Pedregosa2011ScikitLearn}. Gradient boosting models were benchmarked using \texttt{XGBoost}, which is widely regarded for its efficiency and scalability \citep{chen2016xgboost}. Pre-trained transformer models were fine-tuned using \texttt{Hugging Face Transformers}, a library that offers robust implementations of transformer-based architectures \citep{Wolf2019HuggingFace}. Additionally, the \texttt{Sentence Transformers}\cite{reimers-2019-sentence-bert}\cite{reimers-2020-multilingual-sentence-bert} library was used to load embedding models. These tools and libraries are well-regarded in the machine learning community and were chosen for their reliability and performance.

@@

All experiments were conducted using a Kaggle Tesla P100 GPU, which delivered the computational power required to efficiently train and evaluate the models. The Tesla P100 enabled faster processing, allowing us to effectively utilize complex models and perform extensive hyperparameter tuning within a practical timeframe. This setup ensures that the experiments are reproducible and can be replicated by other researchers with access to comparable hardware.

>>>>

This section describes the experimental setup, providing details about data splits, preprocessing steps, hyperparameter tuning, computational resources, and the tools and libraries utilized. The goal is to ensure reproducibility, transparency, and alignment with established best practices in machine learning research.
\begin{figure*}[t]
    \centering
    \includegraphics[width=1.7\columnwidth]{images/confusion_matrices_all_tracks.png}
    \caption{Confusion matrices for each language across Tracks A, B, and C.}
    \label{fig:confusion_matrices}
\end{figure*}

@@

Hyperparameter tuning used Optuna \citep{Akiba2019Optuna} to optimize SVM and XGBoost hyperparameters. Bayesian optimization balanced exploration and exploitation, with configurations assessed on the validation set. The best configuration was selected based on the performance metric.

@@

The model fine-tuning with LoRA, and the training of the MLP and XGBoost models, utilized Binary Cross-Entropy (BCE) as the loss function for Tracks A and C, and Cross-Entropy for Track B, owing to its appropriateness for classification tasks. Meanwhile, the training of the SVM model employed hinge loss.Using LoRA, we fine-tuned the Q, K, and V matrices for feature extractor transformer models, as shown in Table \ref{tab:model_selection}. Given the unbalanced dataset, a weighted loss approach was employed to ensure that the model adequately learned from all classes. Optimization for fine-tuning deep learning models was performed using the \texttt{AdamW} optimizer, which improves upon the standard Adam optimizer by decoupling weight decay and learning rate updates \citep{Loshchilov2019AdamW}. To further enhance training stability and convergence, a cosine annealing learning rate scheduler with restarts\cite{loshchilov2017sgdrstochasticgradientdescent} was employed. This scheduling approach helped adaptively reduce the learning rate over time, facilitating better exploration of the loss landscape and improving generalization. The model was trained for a fixed number of epochs, and early stopping was used to terminate training if the validation performance plateaued, thus avoiding overfitting.
\input{tables/ErrorAnalysis}

@@

The implementation of the experiments utilized several state-of-the-art tools and libraries. The deep learning models were implemented and trained using \texttt{PyTorch} \citep{Paszke2019PyTorch}. For data manipulation and evaluation metrics, \texttt{Scikit-Learn} was employed \citep{Pedregosa2011ScikitLearn}. Gradient boosting models were benchmarked using \texttt{XGBoost} \citep{chen2016xgboost}. Pre-trained transformer models were fine-tuned using \texttt{Hugging Face Transformers} \citep{Wolf2019HuggingFace}. Additionally, the \texttt{Sentence Transformers} library was used to load embedding models \cite{reimers-2019-sentence-bert}\cite{reimers-2020-multilingual-sentence-bert}. These tools and libraries are well-regarded in the machine learning community and were chosen for their reliability and performance.

@@

Experiments used a Kaggle Tesla P100 GPU for efficient model training, evaluation, and hyperparameter tuning, ensuring reproducibility with comparable hardware.
